---
title: "PS-modelReport"
author: "Jarrett Lovelett"
date: "`r format(Sys.time(), '%c')`"
output: html_document
---

```{r setup, include=FALSE}
# memory.limit(size=999999)
# options(mc.cores = 4)
knitr::opts_chunk$set(echo = F, warning = F, message=F)
library(tidyverse)
library(rstan)
library(bayesplot)
library(loo)
load('../output/PS-fitData.rdata')
## should load:
## fit.ps (stan fit object)
## pred.dat.ps (predictions)
## coefs.ps (coefficients) (parameters?)
```

# Process Shift Model Report

## Stan Model
```{r}
print(fit.ps@stanmodel)
```


## Model Fit to Data 
```{r subItemPlot, fig.width = 150, fig.height = 150}
pred.dat.ps %>%
  mutate(strategy = as.factor(strategy)) %>%
  ggplot(aes(x = trial, group=strategy))+
  geom_point(aes(y=RT, color= strategy), size = 6)+
  geom_line(aes(y=pred.RT.ps, group=strategy), size = 4.5, color = 'black', alpha = .8)+
  facet_grid(subject~item)+
  theme(text = element_text(size = 40),
        legend.position='bottom',
        legend.text = element_text(size = 50)
        )
```


## Diagnostics
```{r reformat summary}
sum.ps <- summary(fit.ps)$summary
param <- rownames(sum.ps)
sum.ps <- cbind(param, sum.ps) %>% as_tibble()
sum.ps <- sum.ps %>%
  filter(!startsWith(param, 'y_hat') & 
           !startsWith(param, 'logLik')) %>%
  rowwise() %>%
  mutate(Parameter = case_when(startsWith(param, 'Alg_M') ~ 'Alpha',
                               startsWith(param, 'Ret_B') ~ 'Beta',
                               startsWith(param, 'Ret_T') ~ 'Tau',
                               startsWith(param, 'Ret_R') ~ 'Rate',
                               startsWith(param, 'sigma') ~ 'Sigma',
                               startsWith(param, 'lp_') ~ 'Log Posterior',
                               TRUE ~ 'Transformed/Other')) 

sum.ps <- tibble(Parameter = sum.ps$Parameter, 
                 param = sum.ps$param,
                 as_tibble(
                   lapply(
                   select(sum.ps, -Parameter, -param), 
                   as.numeric)))

# max and min Rhats to display on plots
lims <- c(min(.8,min(sum.ps$Rhat)),
          max(2.2, max(sum.ps$Rhat)))
step <- .2
```

### Rhat distribution
```{r rhat hist}
sum.ps %>% 
  ggplot(aes(x= Rhat)) +
  geom_histogram(bins=1000)+
  geom_vline(aes(xintercept =1), linetype=2)+
  scale_y_continuous("Num. params with Rhat = x", expand=c(0,0))+
  scale_x_continuous(breaks = seq(lims[1],lims[2],step))+
  theme_bw()
```

```{r rhat boxplot}
sum.ps %>%
  ggplot(aes(x = Parameter, y = Rhat))+
  geom_boxplot()+
  geom_hline(aes(yintercept = 1),linetype =2)
```

### Rhats for component parameters

Each spike represents the R-hat value for a particular component parameter. Below, very high values are examined more closely. 
```{r}
plots = sum.ps %>% 
  filter(!Parameter %in% c('Transformed', 'Log Posterior')) %>%
  group_by(Parameter) %>%
  do(plots = ggplot(data=., aes(x = param, y= Rhat)) +
       geom_col()+
       geom_hline(aes(yintercept = 1), linetype =2)+
       ggtitle(unique(.$Parameter))+
       xlab(paste0(.$Parameter, ' Components'))+
       coord_cartesian( 
         ylim = lims)+
       theme(
         axis.text.x = element_blank(),
         axis.ticks.x = element_blank()
       )
     )
lapply(1:length(plots$plots), function(plot){
  plots$plots[[plot]]}) # lol

```

```{r}
z.crit = 1.75 # what defines an extreme value?
```

### Extreme Rhat values (z.score(Rhat) > `r z.crit`)

```{r}
z.score <- function(x){
  (x-mean(x))/sd(x)
}

extreme.rhats <- sum.ps %>% 
  mutate(Rhat.z = z.score(Rhat)) %>%
  filter(abs(Rhat.z) > z.crit) %>%
  arrange(Parameter,param) %>%
  mutate(param = factor(param, levels = param))

extreme.rhats %>%
  ggplot(aes(x = param, y = Rhat))+
  geom_col()+
  geom_hline(aes(yintercept = 1),linetype =2)+
  theme(
         axis.text.x = element_text(size = 5, angle = 45, 
                                    hjust = 1))
knitr::kable(extreme.rhats)
```

```{r examine concerning params}
weird.params <- c('Alg_M_esi[92]')
extract_chains <- function(samps, n_warmup){
  lapply(1:dim(samps)[3], function(i_param){# iterate over parameters
    lapply(1:dim(samps)[2], function(j_chain){ # iterate over chains
      tibble(param = dimnames(samps)$parameters[i_param], # get a dataframe of samples per chain...
             chain = j_chain,
             sample = samps[,j_chain,i_param],
             index=1:length(sample),
             isWarmup = index <= n_warmup)
    }) %>% bind_rows() # ... and staple them together long-wise
  }) # now we have a list (length = n params incldued) containing dataframes of indexed samples by chain
}
samps <- rstan::extract(fit.ps, pars = weird.params, permuted=FALSE, inc_warmup=TRUE)

weird.chains <- extract_chains(samps, n_warmup = 1000)
names(weird.chains) <- weird.params
```

```{r}
lapply(names(weird.chains), function(param) {
  weird.chains[[param]] %>%
    filter(isWarmup==FALSE) %>%
    ggplot(aes(x=index, y = sample),alpha=.5)+
    geom_point()+
    #geom_line()+
    facet_grid(chain ~ .)+
    ggtitle(paste0(param,' samples, by Chain'))
})
```

### Raw Data for Alg_M_esi[92] col (item) = 5, row (subject) = 10
#### (With neighbors for comparison)
```{r}
# I would also look at the raw data corresponding to esi[92] to see if there is a reason why the M parameter might be multimodal?
# Take from original data to get subj-item matrix for lookup

source('../scripts_general/mk_sparse_sub-item_matrix.R')
# reverse lookup subject and item from subj-item vector index
sub_item_esi_92 <- get_sub_item_from_siMat(92, siMat(pred.dat.ps))

subj <- sub_item_esi_92['subject']
it <- sub_item_esi_92['item']
subj <- c(subj-1,subj, subj+1) # will fail stupidly if row or col is max or min... 
it <- c(it-1,it, it+1)
pred.dat.ps %>% 
  filter(subject %in% subj,
         item %in% it ) %>%
  droplevels() %>%
  mutate(strategy= as.factor(strategy)) %>%
  ggplot(aes(x = trial, y = RT, color = strategy))+
  geom_point()+
  geom_line()+
  facet_grid(subject ~ item)
```

Try doing least squares fits to DE data                   